### 2023-05-15 authorshipAndContributions.tex
new version

\section*{Author contributions}

1 - Charles Babbage (Corresponding Author)\\
Computer Scientist, Inventor\\
https://orcid.org/1000-0000-0000-0000 \custombullet~charlesbabbage@example.com\\
Contribution: Conceptualization; Methodology; Writing – Original Draft Preparation

2 - Ada Lovelace \\
Mathematician, Writer\\
https://orcid.org/2000-0000-0000-0000 \custombullet~adalovelace@example.com\\
Contribution: Literature Review, Data Analysis, Writing – Review \& Editing

3 - Pierre Curie\\
Physicist\\
https://orcid.org/1111-2222-3333-6666 \custombullet~pierrecurie@example.com\\
Contribution: Methodology, Data Analysis, Writing – Review \& Editing

4 - Marie Curie\\
Physicist, Chemist\\
https://orcid.org/1111-2222-3333-7777 \custombullet~mariecurie@example.com \\
Contribution: Conceptualization, Methodology, Data Analysis, Writing – Review \& Editing

5 - Grace Hopper\\
Computer Scientist\\
https://orcid.org/1111-2222-3333-8888 \custombullet~gracehopper@example.com\\
Contribution: Conceptualization, Methodology, Writing – Original Draft Preparation

6 - Santos Dumont\\
Inventor, Aviator\\
https://orcid.org/1111-2222-3333-9999 \custombullet~santosdumont@example.com\\
Contribution: Conceptualization, Methodology, Writing – Original Draft Preparation

7 - Nikola Tesla\\
Inventor, Electrical Engineer\\
https://orcid.org/1111-2222-3333-0000 \custombullet~nikolatesla@example.com\\
Contribution: Methodology, Data Analysis, Writing – Review \& Editing

8 - Galileu Galilei\\
Physicist, Mathematician\\
https://orcid.org/1111-2222-3333-1111 \custombullet~galileugalilei@example.com\\
Contribution: Conceptualization, Methodology, Writing – Original Draft Preparation

9 - Jane Goodall\\
Primatologist, Anthropologist\\
https://orcid.org/1111-2222-3333-2222 \custombullet~janegoodall@example.com\\
Contribution: Conceptualization, Writing – Original Draft Preparation, Writing – Review \& Editing

10 - Charles Darwin\\
Biologist, Naturalist\\
https://orcid.org/1111-2222-3333-3333 \custombullet~charlesdarwin@example.com\\
Contribution: Literature Review, Writing – Review \& Editing

11 - Barbara McClintock\\
Geneticist\\
https://orcid.org/1111-2222-3333-4444 \custombullet~barbaramcclintock@example.com\\
Contribution: Methodology, Data Analysis, Writing – Original Draft Preparation